---
title: "Data Analysis Wk3"
format: 
  html:
    self-contained: true
    code-line-numbers: true
  pdf: 
    documentclass: article
custom_title_page: false
filters:
  - lightbox
lightbox: auto
---

---
subtitle: "Data Analysis Week - Session 2"
date: 2023-02-28

author:
  - name: Javier Silva-Valencia
    orcid: 0000-0002-5982-2821
    email: javier.silva@unmsm.edu.pe
    affiliations:
      - name: Instituut Voor Tropische Geneeskunde. Antwerp-Belgium

language: 
  title-block-author-single: "Writer"
  
abstract: Today we are going to create an index/score. For this we have a lot of variables of socioeconomical aspects and we are going to clean and select the ones we are going to use to create the index/score
---

------------------------------------------------------------------------

------------------------------------------------------------------------


### Analysis

Installing packages
```{r}
#install.packages("epiR")
library(epiR)
```



Now we will try to make a multivariate analysis

```{r}
final_dataset = readRDS("final_dataset.RDS")
final_dataset$case <- final_dataset$suffered_vl_since_2nd_survey

```


Brickwall is different to fench wall


#### Mushahar caste
```{r}
table(final_dataset$household_head_subcaste)
```


We'll make an additional factors, 'mushahar' based on 'household_head_subcaste'="MUSHAHAR" 
```{r}
final_dataset$mushahar <- ifelse(final_dataset$household_head_subcaste == "MUSHAHAR", TRUE, FALSE)
table(final_dataset$household_head_subcaste, final_dataset$mushahar)
```

#### Number of people living in each household

First I create a variable 'hh_size' which is the number of rows per FSN:

```{r}
final_dataset$hh_size <- ave(final_dataset$FSN, final_dataset$FSN, FUN = length)
#Calculate the
```


use summaries to see that you have a variable hh_size ranging from 1 to 35
```{r}
summary(final_dataset$hh_size)
```

I also create a variable 'large_hh' based on 'hh_size' 
```{r}
final_dataset$large_hh <- ifelse(final_dataset$hh_size > 7, TRUE, FALSE)
```


#### Now compute a new variable net 3 which is 

No_Mosquito_Net/hh_size. Then recode it, lo:0.333 = FALSE, 0.333:hi = TRUE

```{r}
final_dataset$net3 <- final_dataset$No_Mosquito_Net / final_dataset$hh_size
```
bednet per three person, otherwise FALSE
```{r}
final_dataset$net3 <- final_dataset$net3 >= 0.333
table(final_dataset$net3)
```


#### Check with cc whether net3 and mushahar are associated with case (bivariate)


```{r}
LR_Model3 <- glm(case ~ net3, family=binomial, data=final_dataset)
summary(LR_Model3)
exp(coef(LR_Model3))
exp(confint(LR_Model3))



LR_Model4 <- glm(case ~ mushahar, family=binomial, data=final_dataset)
summary(LR_Model4)
exp(coef(LR_Model4))
round(exp(confint(LR_Model3)),2)
```



# Now make a factor 'riskwall' with three levels: thatched wall, unplastered brick wall and plastered brick wall. 
# check association with case
2 =brick
1 = 
3 = 


```{r}
final_dataset$risk_wall <- ifelse(
  final_dataset$Wall_Material %in% c(6,164,166), 2, ifelse(
    final_dataset$Wall_Material %in% 162:163, 1, ifelse(
      final_dataset$Wall_Material == 165, 3, NA)))

final_dataset$risk_wall <- as.factor(final_dataset$risk_wall)
```




# Ownership of animals
```{r}
final_dataset$own_buf <- ifelse(final_dataset$count_Buf > 0, TRUE, FALSE)
final_dataset$own_cow <- ifelse(final_dataset$count_Cow > 0, TRUE, FALSE)
final_dataset$own_Goa <- ifelse(final_dataset$count_Goa > 0, TRUE, FALSE)
final_dataset$own_Pou <- ifelse(final_dataset$count_Pou > 0, TRUE, FALSE)
```


# Here I create a variable age group
```{r}
final_dataset$agegrp <- cut(final_dataset$member_age, breaks=c(0,9,19,29,39,100))
```


# and a variable 'female'
```{r}
final_dataset$female <- ifelse(final_dataset$member_sex ==3, TRUE, FALSE)
```



# and a variable Damp_floor
```{r}
table(final_dataset$Is_Floor_Damp)
final_dataset$Damp_floor <- ifelse(final_dataset$Is_Floor_Damp == 1, TRUE, FALSE)
```


# Recoding to binary the varaibles Sprayed_20**
```{r}
table(final_dataset$Sprayed_2009)
table(final_dataset$Sprayed_2010)
final_dataset$IRS_09 <- ifelse(final_dataset$Sprayed_2009 > 86, TRUE, FALSE)
final_dataset$IRS_10 <- ifelse(final_dataset$Sprayed_2010 > 86, TRUE, FALSE)
```


#Analysis



# Inputs for my table on the study population:
```{r}
table(final_dataset$agegrp)
table(final_dataset$large_hh)
table(final_dataset$asset_index)
table(final_dataset$Bamboo_Tree)
table(final_dataset$Banana_Tree)
table(final_dataset$case)
table(final_dataset$indoor_Buf)
table(final_dataset$indoor_Cow)
table(final_dataset$indoor_Goa)
table(final_dataset$Is_Floor_Damp)
table(final_dataset$mushahar)
table(final_dataset$Neem_Tree)
table(final_dataset$net3)
table(final_dataset$own_bov)
table(final_dataset$Perm_Water_Body)
table(final_dataset$Rice_Field)
table(final_dataset$risk_wall)
table(final_dataset$Sprayed_2009)
table(final_dataset$Sprayed_2010)
table(final_dataset$female)
table(final_dataset$own_buf)
table(final_dataset$own_cow)
table(final_dataset$own_Goa)
table(final_dataset$own_Pou)
table(final_dataset$Damp_floor)
```


# calculate OR for exposure females
```{r}
LR_Model5 <- glm(case ~ female, family=binomial, data=final_dataset)
summary(LR_Model5)
exp(coef(LR_Model5))
```


# calculate OR for exposure large_hh
```{r}
LR_Model5a <- glm(case ~ large_hh, family=binomial, data=final_dataset)
summary(LR_Model5a)
exp(coef(LR_Model5a))
```


# calculate OR for exposure age group
```{r}
LR_Model6 <- glm(case ~ agegrp, family=binomial, data=final_dataset)
summary(LR_Model6)
exp(coef(LR_Model6))
```


# calculate OR for exposure Mushahar caste
```{r}
LR_Model7 <- glm(case ~ mushahar, family=binomial, data=final_dataset)
summary(LR_Model7)
exp(coef(LR_Model7))
```


# calculate OR for exposure Risk_wall caste
```{r}
LR_Model8 <- glm(case ~ factor(risk_wall), family=binomial, data=final_dataset)
summary(LR_Model8)
exp(coef(LR_Model8))
```


# calculate OR for exposure damp floor
```{r}
LR_Model9 <- glm(case ~ Damp_floor, family=binomial, data=final_dataset)
summary(LR_Model9)
exp(coef(LR_Model9))
```


# calculate OR for exposure sprayed (IRS)
```{r}
LR_Model10 <- glm(case ~ IRS_09, family=binomial, data=final_dataset)
summary(LR_Model10)
exp(coef(LR_Model10))
LR_Model10 <- glm(case ~ IRS_10, family=binomial, data=final_dataset)
summary(LR_Model10)
exp(coef(LR_Model10))
```


# calculate OR for exposure bednet
```{r}
LR_Model11 <- glm(case ~ net3, family=binomial, data=final_dataset)
summary(LR_Model11)
exp(coef(LR_Model11))
```


# calculate OR for exposure ownership of animals
```{r}
LR_Model12 <- glm(case ~ own_Goa, family=binomial, data=final_dataset)
summary(LR_Model12)
exp(coef(LR_Model12))
LR_Model13 <- glm(case ~ own_cow, family=binomial, data=final_dataset)
summary(LR_Model13)
exp(coef(LR_Model13))
LR_Model14 <- glm(case ~ own_buf, family=binomial, data=final_dataset)
summary(LR_Model14)
exp(coef(LR_Model14))
LR_Model15 <- glm(case ~ own_Pou, family=binomial, data=final_dataset)
summary(LR_Model15)
exp(coef(LR_Model15))
```


# calculate OR for exposure trees, rice field, water bodies
```{r}
LR_Model16 <- glm(case ~ Bamboo_Tree, family=binomial, data=final_dataset)
summary(LR_Model16)
exp(coef(LR_Model16))
LR_Model17 <- glm(case ~ Banana_Tree, family=binomial, data=final_dataset)
summary(LR_Model17)
exp(coef(LR_Model17))
LR_Model18 <- glm(case ~ Neem_Tree, family=binomial, data=final_dataset)
summary(LR_Model18)
exp(coef(LR_Model18))
LR_Model19 <- glm(case ~ Perm_Water_Body, family=binomial, data=final_dataset)
summary(LR_Model19)
exp(coef(LR_Model19))
LR_Model20 <- glm(case ~ Rice_Field, family=binomial, data=final_dataset)
summary(LR_Model20)
exp(coef(LR_Model20))
```



# Now I fit a logistic regression model with all factors that were signifcant at p < 0.10. I start by fitting univariate logistic regression
# models for all factors, followed by an LR test

```{r}
GLM.a <- glm(case ~ agegrp, family=binomial(logit), data=final_dataset)
GLM.b <- glm(case ~ Bamboo_Tree, family=binomial(logit), data=final_dataset)
GLM.c <- glm(case ~ Banana_Tree, family=binomial(logit), data=final_dataset)
GLM.d <- glm(case ~ Neem_Tree, family=binomial(logit), data=final_dataset)
GLM.e <- glm(case ~ Perm_Water_Body, family=binomial(logit), data=final_dataset)
GLM.f <- glm(case ~ Rice_Field, family=binomial(logit), data=final_dataset)
GLM.g <- glm(case ~ IRS_09, family=binomial(logit), data=final_dataset)
GLM.h <- glm(case ~ IRS_10, family=binomial(logit), data=final_dataset)
GLM.i <- glm(case ~ mushahar, family=binomial(logit), data=final_dataset)
GLM.j <- glm(case ~ net3, family=binomial(logit), data=final_dataset)
GLM.k <- glm(case ~ risk_wall, family=binomial(logit), data=final_dataset)
GLM.l <- glm(case ~ own_buf, family=binomial(logit), data=final_dataset)
GLM.m <- glm(case ~ own_cow, family=binomial(logit), data=final_dataset)
GLM.n <- glm(case ~ own_Goa, family=binomial(logit), data=final_dataset)
GLM.o <- glm(case ~ own_Pou, family=binomial(logit), data=final_dataset)
GLM.p <- glm(case ~ agegrp, family=binomial(logit), data=final_dataset)
GLM.q <- glm(case ~ female, family=binomial(logit), data=final_dataset)
GLM.r <- glm(case ~ Damp_floor, family=binomial(logit), data=final_dataset)
GLM.s <- glm(case ~ asset_index, family=binomial(logit), data=final_dataset)
```




```{r}
anova(GLM.a, test="Chisq")
anova(GLM.b, test="Chisq")
anova(GLM.c, test="Chisq")
anova(GLM.d, test="Chisq")
anova(GLM.e, test="Chisq")
anova(GLM.f, test="Chisq")
anova(GLM.g, test="Chisq")
anova(GLM.h, test="Chisq")
anova(GLM.i, test="Chisq")
anova(GLM.j, test="Chisq")
anova(GLM.k, test="Chisq")
anova(GLM.l, test="Chisq")
anova(GLM.m, test="Chisq")
anova(GLM.n, test="Chisq")
anova(GLM.o, test="Chisq")
anova(GLM.p, test="Chisq")
anova(GLM.q, test="Chisq")
anova(GLM.r, test="Chisq")
anova(GLM.s, test="Chisq")
```



# Significant at 0.10 are; IRS_09, IRS_10, asset index, banana tree, neem tree, water body, mushahar, net3, own_Goa, own_buf, risk_wall and Damp_Floor.
# I the final model I will consider all except the IRS variables because here I suspect reversed causality

```{r}
GLM.1 <- glm(case ~ asset_index + Banana_Tree + Neem_Tree + mushahar + Perm_Water_Body + net3 + own_Goa + own_buf + risk_wall + Damp_floor, 
             family=binomial(logit), data=final_dataset)
summary(GLM.1)
exp(coef(GLM.1))  # Exponentiated coefficients ("odds ratios")
```


# Weakest is neem tree with a p-value of 0.949, so I drop it

```{r}
GLM.2 <- glm(case ~ asset_index + Banana_Tree + mushahar + Perm_Water_Body + net3 + own_Goa + own_buf + risk_wall + Damp_floor, 
             family=binomial(logit), data=final_dataset)
summary(GLM.2)
exp(coef(GLM.2))  # Exponentiated coefficients ("odds ratios")
```


```{r}
anova(GLM.1, GLM.2, test="Chisq")
```


# weakest is now damp floor with a p-value of 0.944, so I drop it

```{r}
GLM.3 <- glm(case ~ asset_index + Banana_Tree + mushahar + Perm_Water_Body + net3 + own_Goa + own_buf + risk_wall, 
             family=binomial(logit), data=final_dataset)
summary(GLM.3)
exp(coef(GLM.3))  # Exponentiated coefficients ("odds ratios")
```

```{r}
anova(GLM.2, GLM.3, test="Chisq")
```


# weakest is now net3, p = 0.247 so I drop it

```{r}
GLM.4 <- glm(case ~ asset_index + Banana_Tree + mushahar + Perm_Water_Body + own_Goa + own_buf + risk_wall, 
             family=binomial(logit), data=final_dataset)
summary(GLM.4)
exp(coef(GLM.4))  # Exponentiated coefficients ("odds ratios")
```

```{r}
anova(GLM.3, GLM.4, test="Chisq")
```


# weakest is now banana tree, p = 0.244107, so I drop it

```{r}
GLM.5 <- glm(case ~ asset_index + mushahar + Perm_Water_Body + own_Goa + own_buf + risk_wall, 
             family=binomial(logit), data=final_dataset)
summary(GLM.5)
exp(coef(GLM.5))  # Exponentiated coefficients ("odds ratios")
```


```{r}
anova(GLM.4, GLM.5, test="Chisq")
```


# weakest is now permanent water body with a p-value of 0.12, si I drop it

```{r}
GLM.5 <- glm(case ~ asset_index + mushahar + own_Goa + own_buf + risk_wall, 
             family=binomial(logit), data=final_dataset)
summary(GLM.5)
exp(coef(GLM.5))  # Exponentiated coefficients ("odds ratios")
```


```{r}
anova(GLM.4, GLM.5, test="Chisq")
```


# weakest is now own_buf with a p-value of 0.078, so I drop it

```{r}
GLM.6 <- glm(case ~ asset_index + mushahar + own_Goa + risk_wall, 
             family=binomial(logit), data=final_dataset)
summary(GLM.6)
exp(coef(GLM.6))  # Exponentiated coefficients ("odds ratios")
```

```{r}
anova(GLM.5, GLM.6, test="Chisq")
```


# weakest is now own_Goa, with a p-value of 0.059, so I drop it

```{r}
GLM.7 <- glm(case ~ asset_index + mushahar + risk_wall, 
             family=binomial(logit), data=final_dataset)
summary(GLM.7)
exp(coef(GLM.7))  # Exponentiated coefficients ("odds ratios")
```

```{r}
anova(GLM.6, GLM.7, test="Chisq")
```


# the LR test is non significant, p = 0.06 so I can indeed drop own_Goa



# Now all factors are significant for at least one level. Asset index is weakest, I will just try without it

```{r}
GLM.8 <- glm(case ~ mushahar + risk_wall, family=binomial(logit), data=final_dataset)
summary(GLM.8)
```

```{r}
exp(coef(GLM.8))  # Exponentiated coefficients ("odds ratios")
```


# Clearly the model with asset index is significantly better than the model without, p= 0.0008.
```{r}
anova(GLM.7, GLM.8, test="Chisq")
```


# Next I fit an interaction between risk wall and asset index
```{r}
GLM.9 <- glm(case ~ mushahar + risk_wall + asset_index  +asset_index*risk_wall, family=binomial(logit), data=final_dataset)
summary(GLM.9)
exp(coef(GLM.9))  # Exponentiated coefficients ("odds ratios")
```


# And I do the likelihood ratio testing

```{r}
anova(GLM.7, GLM.9, test="Chisq")
```


# p-value of LR test is 0.72, so model with interaction is not significnatly better

# Next I try interaction between Mushahar and asset index
```{r}
GLM.10 <- glm(case ~ mushahar + risk_wall + asset_index + asset_index * mushahar, family=binomial(logit), data=final_dataset)
summary(GLM.10)
exp(coef(GLM.10))  # Exponentiated coefficients ("odds ratios")
```


# LR test has a p-value of 0.09, so complex model not significantly better
```{r}
anova(GLM.7, GLM.10, test="Chisq")
```


# Finally I test the interaction between Mushahar and asset index

```{r}
GLM.11 <- glm(case ~ mushahar + risk_wall + asset_index + risk_wall* mushahar, family=binomial(logit), data=final_dataset)
summary(GLM.11)
exp(coef(GLM.11))  # Exponentiated coefficients ("odds ratios")
```


# LR test has a p-vaule of 0.69, so no indication for an interaction
```{r}
anova(GLM.11, GLM.7, test="Chisq")
```


# So my final model remains model 7 with asset index, mushahar and risk wall

```{r}
GLM.7 <- glm(case ~ mushahar + risk_wall + asset_index, family=binomial(logit), data=final_dataset)
summary(GLM.7)
exp(coef(GLM.7))  # Exponentiated coefficients ("odds ratios")
exp(confint(GLM.7))
```

Final model: GLM.7


# You could consider recoding asset index to a binary varibale poverty, 1:4 = TRUE, 5 = FALSE
# The interaction terms remain non significant. 
If you really want to check for interaction is better to have two categories

Creation variable poverty
```{r}
final_dataset$poverty <- ifelse(final_dataset$asset_index %in% 1:4, TRUE, FALSE)
```


```{r}
GLM.14 <- glm(case ~ mushahar + risk_wall + poverty, family=binomial(logit), data=final_dataset)
summary(GLM.14)
exp(coef(GLM.14))  # Exponentiated coefficients ("odds ratios")
exp(confint(GLM.14))
```


```{r}
GLM.15 <- glm(case ~ mushahar + risk_wall + poverty+ poverty*mushahar, family=binomial(logit), data=final_dataset)
summary(GLM.15)
exp(coef(GLM.15))  # Exponentiated coefficients ("odds ratios")
```

```{r}
anova(GLM.15, GLM.14, test="Chisq")
exp(confint(GLM.14))
```


```{r}

GLM.16 <- glm(case ~ mushahar + risk_wall + poverty + poverty * risk_wall, family=binomial(logit), data=final_dataset)
summary(GLM.16)
exp(coef(GLM.16))  # Exponentiated coefficients ("odds ratios")
```

```{r}
anova(GLM.16, GLM.14, test="Chisq")
```



```{r}
GLM.17 <- glm(case ~ mushahar + risk_wall + poverty + mushahar * risk_wall, family=binomial(logit), data=final_dataset)
summary(GLM.17)
exp(coef(GLM.17))  # Exponentiated coefficients ("odds ratios")
```

```{r}
anova(GLM.17, GLM.14, test="Chisq")
```


# So then my final model here is model 14: case = poverty + risk_wall + mushahar, but since poverty did not have any interactions we go back to model 7. Model 7 is the final model.

```{r}
GLM.7 <- glm(case ~ mushahar + risk_wall + asset_index, family=binomial(logit), data=final_dataset)
summary(GLM.7)
exp(coef(GLM.7))  # Exponentiated coefficients ("odds ratios")
exp(confint(GLM.7))
```


# odds ratios: mushahar 7.2 (4.4-9.9), brick wall vs thatched wall 0.7 (0.5-0.9), plastered brick wall vs. thatched wall 0.3 (0.2-0.5)
# and assetindex 1.7(1.2-2.6), 0.9(0.6-1.4), 1.1(0.7-1.7), 0.6(0.3-1.1)

# At the end I export my final dataset to csv so I can use it in other programs
write.table(final_dataset, "final_dataset.csv", sep=",", col.names=TRUE, 
            row.names=FALSE, quote=TRUE, na="NA")


#The poorest of the poor: bihar viscerlar leshmaniasis
#Visceral Leshmaniasis in Bihar 
